---
title: "Modele_semestriel_3"
output: html_document
date: "2023-09-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(openxlsx)
library(car)
library(shiny)
library(MASS)
library(tidyverse)
library(rstan)
library(ggpubr)
library(fitdistrplus)
library(nortest)
library(fitur)
library(actuar)
library(moments)
library(AID)
library(FAdist)
library(rmarkdown)
library(tinytex)
library(numDeriv)
library(rgl)
library(lmtest)
library(pander)
library(sae)
library(zoo)
library(lmtest) 
library(forecast)
library(fBasics)
library(GGally)
library(broom)
options(scipen = 999,digits=3)
 b8<- read.xlsx("C:/Users/wassi/OneDrive/Bureau/BilanSem2.xlsx")
 b8$date<-as.Date(b8$date,origin = "1899-12-30")
 z<- read.xlsx("C:/Users/wassi/OneDrive/Bureau/BilanSem_trans.xlsx")
```

```{r}
# ggpairs(b8[,2:21])
# pairs(b8[,2:ncol(b8)])
# cor(b8$ac3,b8$pa3)

b9<-cbind(AC12=b8$ac1+b8$ac2,AC3=b8$ac3,AC45=b8$ac4+b8$ac5,AC6=b8$ac6,AC7=b8$ac7,PA12=b8$pa1+b8$pa2,PA3=b8$pa3,PA4=b8$pa4,PA5=b8$pa5,TP=b8$tp,CP1=b8$cp1,CP2=b8$cp2,CP4=b8$cp4,CP56= na.omit(as.numeric(z[15,])),TCP=b8$tcp,TB=b8$tb)
b9<-data.frame(b9)
cm<-cor(b9,method="pearson")
view(cm)

```
#### A partir de la matrice de corrélation des variables du bilan, on peut obtenir une première idée sur la dépendance de chaque variable pour construire un modèle de départ.

#### Check model summary: eliminate insignificant predictors (high p-values) and look for R-adjusted evolution.
#### Check for multicollinearity with vif() function;#In general, a VIF greater than 10 suggests a problematic amount of collinearity.
#### Carry on until fitting the best model (highest possible  R-adjusted )
#### Step() function tries multiple lm combos, we might also select the direction: backward,forward or both 
#### We will eliminate sum variables TP, TB, TCP to avoid colinearity issues.

## AC12

```{r}
#lmac12<-lm(formula=b9$AC12~.,data=b9) with all variables

#Essai 1:

lmac12<-lm(formula=b9$AC12~b9$AC45+ b9$AC6+ b9$AC7+b9$PA3+ b9$PA4 + b9$PA5 + b9$CP1 + b9$CP2 + b9$CP56,data=b9)
summary(lmac12)
vif(lmac12)
print(summary(lmac12)$adj.r.squared)
#Essai 2:
lmac12<-lm(formula = b9$AC12 ~ b9$AC45 + b9$CP1 + b9$CP2+b9$CP56+b9$PA4, data = b9)
summary(lmac12)
vif(lmac12)
print(summary(lmac12)$adj.r.squared)
#Essai3:
lmac12<-step(lmac12, direction="both")
summary(lmac12)
vif(lmac12)
print(summary(lmac12)$adj.r.squared)
#predict.lm(lmac12,data.frame(A45=555555,CP1=789,CP2=897846,CP56=78852)) a revoir
```


#### Conclusion: AC12 ~ -1.045*AC45 + 37.408 *CP1 + 1.707*CP2 + 1.050 *CP56 - 3645039.824


## AC3

```{r}
#Essai 1:

lmac3<-lm(formula=b9$AC3~b9$AC45+ b9$AC6+ b9$AC7+b9$PA3+ b9$PA12 + b9$PA5 + b9$CP1 + b9$CP2 + b9$CP56,data=b9)
summary(lmac3)
vif(lmac3)
print(summary(lmac3)$adj.r.squared)
#Essai 2:
lmac3<-lm(formula=b9$AC3~b9$AC45+ b9$AC7+ b9$PA3+ b9$CP1+b9$CP2,data=b9)
summary(lmac3)
vif(lmac3)
print(summary(lmac3)$adj.r.squared)
#Essai3:
lmac3<-step(lmac3, direction="both")
summary(lmac3)
vif(lmac3)
print(summary(lmac3)$adj.r.squared)
```

#### Conclusion: AC3 ~ 0.913*AC45 -3.397*AC7  -1.356 *CP2 + 0.566*PA3




## AC45

```{r}
#Essai 1:

lmac45<-lm(formula=b9$AC45~b9$AC12+ b9$AC3+ b9$AC6+ b9$AC7+b9$PA3+b9$PA4+ b9$PA12 + b9$PA5 + b9$CP1 + b9$CP2 + b9$CP56,data=b9)
summary(lmac45)
vif(lmac45)
print(summary(lmac3)$adj.r.squared)

#Essai 2:

lmac45<-lm(formula=b9$AC45~b9$AC12+ b9$AC3+ b9$AC7+b9$PA4+ b9$PA12 + b9$PA5 + b9$CP1 + b9$CP56,data=b9)
summary(lmac45)
vif(lmac45)
print(summary(lmac3)$adj.r.squared)
#Essai3:
lmac45<-step(lmac45, direction="both")
summary(lmac45)
vif(lmac45)
print(summary(lmac45)$adj.r.squared)
```

#### Conclusion: AC3 ~ 0.6019*AC3 +3.0873*AC7+34.5257*CP1 + 0.8305*PA4 -1.1376*CP56-7422762.4156



## A6

```{r}
#Essai 1:

lmac6<-lm(formula=b9$AC6~b9$AC12+ b9$AC3+ b9$AC7+b9$PA3+b9$PA4 + b9$PA5 + b9$CP1 + b9$CP2 + b9$CP56,data=b9)
summary(lmac6)
vif(lmac6)
print(summary(lmac6)$adj.r.squared)

#Essai 2:

lmac6<-lm(formula=b9$AC6~b9$AC7+ b9$CP2 + b9$CP56,data=b9)
summary(lmac6)
vif(lmac6)
print(summary(lmac6)$adj.r.squared)
#Essai3:
lmac6<-step(lmac6, direction="both")
summary(lmac6)
vif(lmac6)
print(summary(lmac6)$adj.r.squared)
```

#### Conclusion: AC6 ~ -0.570*AC7+ 0.4*CP2 + 0.568*CP56-48296.988


## AC7

```{r}
#Essai 1:

lmac7<-lm(formula=b9$AC7~b9$AC12+ b9$AC3+b9$AC45+ b9$AC6+b9$PA3+b9$PA12+ b9$PA5 + b9$CP1 + b9$CP2 + b9$CP56,data=b9)
summary(lmac7)
vif(lmac7)
print(summary(lmac7)$adj.r.squared)
#Essai 2:
lmac7<-lm(formula=b9$AC7~b9$AC12+ b9$AC3+ b9$AC6+b9$PA3+b9$PA12,data=b9)
summary(lmac7)
vif(lmac7)
print(summary(lmac7)$adj.r.squared)
#Essai3:
lmac3<-step(lmac7, direction="both")
summary(lmac7)
vif(lmac7)
print(summary(lmac7)$adj.r.squared)
```


#### Conclusion: AC7 ~ -0.1481*AC12 -0.2884*AC3-0.3828*AC6+0.19 *PA12 +0.2604*PA3





## PA12

```{r}
#Essai 1:

lmpa12<-lm(formula=b9$PA12~b9$AC45+b9$AC7+b9$AC3+b9$AC6+ b9$CP1 + b9$CP2,data=b9)
summary(lmpa12)
vif(lmpa12)
print(summary(lmpa12)$adj.r.squared)

#Essai 2:

lmpa12<-lm(formula=b9$PA12~b9$AC45+b9$AC7+b9$AC6+ b9$CP1 + b9$CP2,data=b9)
summary(lmpa12)
vif(lmpa12)
print(summary(lmpa12)$adj.r.squared)
#Essai3:
lmpa12<-step(lmpa12, direction="both")
summary(lmpa12)
vif(lmpa12)
print(summary(lmpa12)$adj.r.squared)
```

#### Conclusion: PA12 ~  0.518 *AC45+ 44.514*CP1-1.410*CP2--4606954.531




## PA3

```{r}
#Essai 1:

lmpa3<-lm(formula=b9$PA3~ b9$AC6+ b9$PA5 + b9$CP1 ,data=b9)
summary(lmpa3)
vif(lmpa3)
print(summary(lmpa3)$adj.r.squared)

#Essai2:
lmpa3<-step(lmpa3, direction="both")
summary(lmpa3)
vif(lmpa3)
print(summary(lmpa3)$adj.r.squared)
```

#### Conclusion: pa3 ~ + b9$AC7*AC6 -6.03*CP1 + 3.05*PA5 + 4507277.69



## PA4

```{r}
#Essai 1:

lmpa4<-lm(formula=b9$PA4~b9$AC12+ b9$AC45+ b9$AC6+b9$PA12+b9$PA3,data=b9)
summary(lmpa4)
vif(lmpa4)
print(summary(lmpa4)$adj.r.squared)

#Essai2:
lmpa4<-step(lmpa4, direction="both")
summary(lmpa4)
vif(lmpa4)
print(summary(lmpa4)$adj.r.squared)
```

#### Very low R²: PA4 has a weak linear correlation with the rest of the variables.


## PA5

```{r}
#Essai 1:

lmpa5<-lm(formula=b9$PA5~b9$AC45+ b9$AC3+ b9$AC6+b9$PA3+ b9$CP2,data=b9)
summary(lmpa5)
vif(lmpa5)
print(summary(lmpa5)$adj.r.squared)

#Essai2:
lmpa5<-step(lmpa5, direction="both")
summary(lmpa5)
vif(lmpa5)
print(summary(lmpa5)$adj.r.squared)
```

#### Conclusion: PA5 ~ -0.6410*AC6+ 0.1182*PA3 -338253.4621




## CP1

```{r}
#Essai 1:

lmcp1<-lm(formula=b9$CP1~b9$AC12+ b9$AC3+ b9$AC7+ b9$AC6+ b9$AC45+ b9$PA12+ b9$PA3 + b9$CP2 + b9$CP56,data=b9)
summary(lmcp1)
vif(lmcp1)
print(summary(lmcp1)$adj.r.squared)

#Essai 2:

lmcp1<-lm(formula=b9$CP1~ b9$AC7+ b9$AC6+ b9$PA12 + b9$CP2 + b9$CP56,data=b9)
summary(lmcp1)
vif(lmcp1)
print(summary(lmcp1)$adj.r.squared)
#Essai3:
lmcp1<-step(lmcp1, direction="both")
summary(lmcp1)
vif(lmcp1)
print(summary(lmcp1)$adj.r.squared)
```

#### Conclusion: CP1 ~ 0.03306*AC7+ 0.00999*CP2 + 0.00634*PA12 + 109975.33219 (insignificant :CP1(le capital)=cst )




## CP2

```{r}
#Essai 1:

lmcp2<-lm(formula=b9$CP2~b9$AC12+ b9$AC3+ b9$AC6+ b9$AC45+ b9$AC7+ b9$PA12+ b9$PA3+ b9$PA5  +b9$CP56,data=b9)
summary(lmcp2)
vif(lmcp2)
print(summary(lmcp2)$adj.r.squared)

#Essai 2:

lmcp2<-lm(formula=b9$CP2~b9$AC12+ b9$AC6+ b9$AC45 +b9$CP56,data=b9)
summary(lmcp2)
vif(lmcp2)
print(summary(lmcp2)$adj.r.squared)
#Essai3:
lmcp2<-step(lmcp2, direction="both")
summary(lmcp2)
vif(lmcp2)
print(summary(lmcp2)$adj.r.squared)
```

#### Conclusion: CP2 ~ 1.2185*AC6 + 0.2678*AC45 + 0.2661*AC12 -0.8918 *CP56-94719.9720


## CP4 =cst.

## CP56

```{r}
#Essai 1:

lmcp56<-lm(formula=b9$CP56~b9$AC12+ b9$AC6+ b9$AC45+ b9$AC7+ b9$PA3+b9$CP2+b9$CP1,data=b9)
summary(lmcp56)
vif(lmcp56)
print(summary(lmcp56)$adj.r.squared)

#Essai 2:

lmcp56<-lm(formula=b9$CP56~b9$AC12+ b9$AC6+ b9$AC7+b9$CP2,data=b9)
summary(lmcp56)
vif(lmcp56)
print(summary(lmcp56)$adj.r.squared)
#Essai3:
lmcp56<-step(lmcp56, direction="both")
summary(lmcp56)
vif(lmcp56)
print(summary(lmcp56)$adj.r.squared)
```

#### Conclusion: CP56 ~ 0.6664*AC7+ 0.0597*AC12+0.8606*AC6 -0.3775*cp2+48358.8010


```{r}
ff<-data.frame()
ff<-cbind(b9$AC12,predict.lm(lmac12),b9$AC3,predict.lm(lmac3),b9$AC45,predict.lm(lmac45),b9$AC6,predict.lm(lmac6),b9$AC7,predict.lm(lmac7),b9$PA12,predict.lm(lmpa12),b9$PA3,predict.lm(lmpa3),b9$PA4,predict.lm(lmpa4),b9$PA5,predict.lm(lmpa5),b9$CP2,predict.lm(lmcp2),b9$CP56,predict.lm(lmcp56))
rr <- c("juin_2013","dec_2013","juin_2014","dec_2014","juin_2015","dec_2015","juin_2016","dec_2016","juin_2017","dec_2017","juin_2018","dec_2018","juin_2019","dec_2019","juin_2020","dec_2020","juin_2021","dec_2021","juin_2022","dec_2022")
cc<-c("AC12","AC12_fit","AC3","AC3_fit","AC45","AC45-fit","AC6","AC6_fit","AC7","AC7_fit","PA12","PA12_fit","PA3","PA3_fit","PA4","PA4_fit","PA5","PA5_fit","CP2","CP2_fit","CP56","CP56_fit")

rownames(ff) <- rr

knitr::kable(ff, col.names = cc, row.names = TRUE,caption = "Semestrial balance Sheet Comparison: Actual vs Fitted with lm")
pander(ff) 

```




